Meta-analytic approaches and effect sizes to account for 'nuisance heterogeneity' in comparative physiology

### 2021-09-21 MS/ms.Rmd
Justed updated akwnow
Added in Laguenesse 2010

Meta-analysis is always concerned with effect heterogeneity; in other words, the factors that drive differences in the direction and magnitude of effects within and across studies [@Nakagawa2017; @Gurevitch2018; @Cooperetal2019; @Borenstein2019]. The concept of heterogeneity is vitally important because it tells us how general our findings are likely to be and how much of the variance we see is the result of real biology or methodological differences (after meta-analytically accounting for sampling variance) [@Nakagawa2017; @Borenstein2019].

@@

Besides sampling variance, all the sources of heterogeneity discussed above contribute (to varying extents) to the 'apples and oranges' problem often touted as compromising the reliability of meta-analyses. Combining standardised effects, while ignoring these sources of heterogeneity can indeed weaken the reliability of a meta-analysis. However, many argue that heterogeneity is just what we are interested in synthesizing [@Cooperetal2019; @Borenstein2019; @Gurevitch2018; @Glass2015]. In the words of Gene Glass, the founder of meta-analysis: "Of course it [meta-analysis] mixes apples and oranges; in the study of fruit nothing else is sensible; comparing apples and oranges is the only endeavor worthy of true scientists; comparing apples to apples is trivial" [p.g., 224, @Glass2015]. Whether comparing apples and oranges is a good idea simply boils down to the specific question of interest, the population one wishes to make conclusions about and how heterogeneous effects actually are. If we are interested in generalising only to apples, then mixing fruit will not be ideal. However, if we want to understand why different 'fruit' vary in their properties, then we need to consider apples and oranges together [@Borenstein2019].

@@

We would like to thank the Editorial board (Craig Franklin, Michaela Handel and Charlotte Rutledge) for inviting us to contribute to this special issue along with feedback on previous drafts. We would also like to thank Megan Head for comments on a previous draft. DWAN, SMD, SN, ML and REO were supported by ARC Discovery Projects (DP210101152 to DWAN, DE180100202 to SMD, and DP200100367 to SN, ML and DWAN). PP and SB were supported by a UNSW Scientia PhD Scholarship.

>>>>

Meta-analysis is always concerned with effect heterogeneity; in other words, the factors that drive differences in the direction and magnitude of effects within and across studies [@Nakagawa2017; @Gurevitch2018; @Cooperetal2019; @Borenstein2019; @Lag2010]. The concept of heterogeneity is vitally important because it tells us how general our findings are likely to be and how much of the variance we see is the result of real biology or methodological differences (after meta-analytically accounting for sampling variance) [@Nakagawa2017; @Borenstein2019; @Lag2010].

@@

Besides sampling variance, all the sources of heterogeneity discussed above contribute (to varying extents) to the 'apples and oranges' problem often touted as compromising the reliability of meta-analyses. Combining standardised effects, while ignoring these sources of heterogeneity can indeed weaken the reliability of a meta-analysis. However, many argue that heterogeneity is just what we are interested in synthesizing [@Cooperetal2019; @Borenstein2019; @Gurevitch2018; @Glass2015; @Lag2010]. In the words of Gene Glass, the founder of meta-analysis: "Of course it [meta-analysis] mixes apples and oranges; in the study of fruit nothing else is sensible; comparing apples and oranges is the only endeavor worthy of true scientists; comparing apples to apples is trivial" [p.g., 224, @Glass2015]. Whether comparing apples and oranges is a good idea simply boils down to the specific question of interest, the population one wishes to make conclusions about and how heterogeneous effects actually are. If we are interested in generalising only to apples, then mixing fruit will not be ideal. However, if we want to understand why different 'fruit' vary in their properties, then we need to consider apples and oranges together [@Borenstein2019].

@@

We would like to thank the Editorial board (Craig Franklin, Michaela Handel and Charlotte Rutledge) for inviting us to contribute to this special issue along with feedback on previous drafts. We would also like to thank Megan Head and two anonymous reviewers for helpful comments on previous drafts. DWAN, SMD, SN, ML and REO were supported by ARC Discovery Projects (DP210101152 to DWAN, DE180100202 to SMD, and DP200100367 to SN, ML and DWAN). PP and SB were supported by a UNSW Scientia PhD Scholarship.
